---
title: "Time Series Analysis of Daily Wheat and Corn Futures Prices"
author: "Anthony Le" 
date: "Dec 5, 2024" 
editor: visual 
format:         
  html:
    code-summary: "Show the code" 
    warning : false 
    eval: true
    self-contained: true 
    toc: true 
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(astsa) 
library(tstools)
library(xts)
library(mosaic)
library(dplyr)
library(car)
library(Stat2Data)
library(dynlm)
library(nlme)
library(AER)
library(forecast)
library(mgcv)
library(tseries)
library(lmtest)
library(fBasics)
library(leaps)

library(urca)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(gt)
library(kableExtra)
library(modelsummary)
library(performance)
library(utils)
library(scales)
library(gridExtra)
library(DHARMa)
library(nortest)
library(moments)
library(see)
library(patchwork)
library(gganimate)
library(gapminder)
library(gifski)
library(zoo)
library(fGarch) 
library(arfima)
library(tsDyn) 
library(lme4)
library(tsbox)


library(RColorBrewer)
library(TTR)
library(quantmod)
library(PerformanceAnalytics)
library(Quandl)
library(fGarch)
library(parallel)
library(rugarch)




```

# I. Introduction

# II. Methods

## 1. Working Data

### a. Data Acquisition & Wrangling

```{r}

# Data Imports & Wrangling 
wheat <- read.csv("data/wheat_futures.csv")
corn <- read.csv("data/corn_futures.csv")

wheatvol <- read.csv("data/wheat_volume.csv")

# Fill in N/A's for wheat volume
merged_data <- merge(wheat, wheatvol, by = "Date", all.x = TRUE)
merged_data$Vol.[is.na(merged_data$Vol.)] <- merged_data$Volume[is.na(merged_data$Vol.)]
wheat <- merged_data[, !names(merged_data) %in% "Volume"]

wheat <- wheat %>%
  mutate(Date = as.Date(wheat$Date)) 

corn <- corn %>%
  mutate(Date = as.Date(corn$Date))

total <- merge(wheat,corn,by="Date")%>% 
  rename('Wheat' = 'Price.x',
         'wheat_vol' = 'Vol..x',
         'Corn' = 'Price.y',
         'corn_vol' = 'Vol..y') %>%
  mutate(wheat_vol = na.approx(wheat_vol))%>%
  mutate(corn_vol = na.approx(corn_vol))%>%
  mutate("Month" = month(Date)) %>%
  mutate("Year" = year(Date))

```

### b. Potential Biases & Assumptions

## 2. Time Series Analysis

### a. Assumption & Validation

### b. Function of Time Models

### c. ARIMA, ARFIMA, SARIMA, and ARIMAX Models

### d. Spectral Analysis

### e. GARCH Models

### f. Markov Switching AR/Threshold Models

### h. Bootstraping

### i. Cross Validation

### j. Self-defining Functions

```{r, include=FALSE}
# Self-Defining Functions

    # Stationarity Tests
    check_stationarity <- function(data) {
      adf_result <- adf.test(data)
      pp_result <- pp.test(data)
      kpss_result <- kpss.test(data)
      urers_result <- ur.ers(data, type = "DF-GLS", model = "constant")
      results <- data.frame(
        Test = c("ADF Test", "Phillips-Perron Test", "KPSS Test", "ERS Test"),
        `Test Statistic` = c(round(adf_result[["statistic"]][["Dickey-Fuller"]], 3),
                             round(pp_result[["statistic"]][["Dickey-Fuller Z(alpha)"]], 3),
                             round(kpss_result[["statistic"]][["KPSS Level"]], 3),
                             round(urers_result@teststat, 3)),
        `P-Value` = c(round(adf_result[["p.value"]],2),
                      round(pp_result[["p.value"]],2),
                      round(kpss_result[["p.value"]],2),
                      "N/A"),
        Interpretation = c(
          ifelse(adf_result[["p.value"]] < 0.05, "Stationary", "Non-Stationary"),
          ifelse(pp_result[["p.value"]] < 0.05, "Stationary", "Non-Stationary"),
          ifelse(kpss_result[["p.value"]] < 0.05, "Non-Stationary", "Stationary"),
          ifelse(urers_result@teststat < urers_result@cval[2], "Stationary", "Non-Stationary")))
      kable(results, align = 'c') %>%
        kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
        add_header_above(c("Stationarity Test Summary" = 4))
    }

  # Time Series Plotting
    tsgplot <- function(data, X, Y, xlab, ylab, timebreak, name) {
    ggplot(data, aes(x = X, y = Y)) +
      geom_line(color = "darkblue", size = 1) +
      ggtitle(name) +
      xlab(xlab) +
      ylab(ylab) +
      scale_x_date(date_labels = "%b %Y", date_breaks = timebreak)+
      theme_light()+
      scale_y_continuous(labels = scales::label_number(scale_cut = cut_short_scale()))
    }




    # ACF/PACF/First-Last Third Plotting
    four_plots <- function(data) {
      n <- length(data)
      first_third <- data[1:floor(n / 3)]
      last_third <- data[floor(2 * n / 3):n]
      par(mfrow = c(2, 2))
      acf(data, main = "ACF of Whole Data", lag.max =30)
      pacf(data, main = "PACF of Whole Data", lag.max =30)
      acf(first_third, main = "ACF of First Third")
      acf(last_third, main = "ACF of Last Third")
      par(mfrow = c(1, 1))
    }


    # ACF/PACF Plotting
    two_plots <- function(data, lag =50) {
      par(mfrow = c(2, 1))
      acf(data, lag.max =lag)
      pacf(data, lag.max =lag)
      par(mfrow = c(1, 1))
    }

    # Differencing Plotting
    diff_plots <- function(data, name = "Time Series Plot", lag = 30) {
      layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
      astsa::tsplot(data, main= name)
      abline(h = mean(data, na.rm = TRUE), col = "red", lty = 2)
      acf(data, lag.max =lag)
      pacf(data, lag.max =lag)
      par(mfrow = c(1,1))
    }

    xts_plots <- function(data, name = "Time Series Plot", lag = 30) {
      layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
      astsa::tsplot(ts_ts(data), main= name)
      abline(h = mean(ts_ts(data), na.rm = TRUE), col = "red", lty = 2)
      acf(data, lag.max =lag)
      pacf(data, lag.max =lag)
      par(mfrow = c(1,1))
    }



    # Model Summary
    model_summary_table <- function(models_list) {
      model_names <- names(models_list)
      if (is.null(model_names)) {
        model_names <- paste("Model", seq_along(models_list)) }
      summary_df <- data.frame(Model = character(),
                               `Adjusted  $R^2$ ` = numeric(),
                               AIC = numeric(),
                               BIC = numeric(),
                               stringsAsFactors = FALSE)
      for (i in seq_along(models_list)) {
        model <- models_list[[i]]
        model_summary <- summary(model)
        summary_df[i, ] <- c(
          model_names[i],
          round(model_summary[["adj.r.squared"]], 2),
          round(AIC(model), 2),
          round(BIC(model), 2)) }
      kable(summary_df, align = 'c', col.names = c("Model", "Adjusted R²", "AIC", "BIC")) %>%
        kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
        add_header_above(c("Model Summary" = 4))
    }


    # GARCH Model Summary
    garch_table <- function(models_list) {
      model_names <- names(models_list)
      if (is.null(model_names)) {
        model_names <- paste("Model", seq_along(models_list)) }
      summary_df <- data.frame(Model = character(),
                               AIC = numeric(),
                               BIC = numeric(),
                               stringsAsFactors = FALSE)
      for (i in seq_along(models_list)) {
        model <- models_list[[i]]
        summary_df[i, ] <- c(
          model_names[i],
          round(model@fit[["ics"]][["AIC"]], 3),
          round(model@fit[["ics"]][["BIC"]], 3)) }
      kable(summary_df, align = 'c', col.names = c("Model",  "AIC", "BIC")) %>%
        kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
        add_header_above(c("GARCH Summary" = 3))
    }


    # SARIMA Model Summary
    sarima_summary_table <- function(models_list) {
      model_names <- names(models_list)
      if (is.null(model_names)) {
        model_names <- paste("Model", seq_along(models_list)) }
      summary_df <- data.frame(Model = character(),
                               AIC = numeric(),
                               BIC = numeric(),
                               stringsAsFactors = FALSE)
      for (i in seq_along(models_list)) {
        model <- models_list[[i]]
        model_summary <- summary(model)
        summary_df[i, ] <- c(
          model_names[i],
          round(AIC(model), 2),
          round(BIC(model), 2)) }
      kable(summary_df, align = 'c', col.names = c("Model",  "AIC", "BIC")) %>%
        kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
        add_header_above(c("Model Summary" = 3))
    }



    # Normality Tests Function
    check_normality_residuals <- function(model) {
      residuals <- residuals(model) 
      ad_result <- ad.test(residuals)
      jb_result <- jarque.bera.test(residuals)
      lillie_result <- lillie.test(residuals)
      skew <- skewness(residuals)
      kurt <- kurtosis(residuals)
      # Summary table
      results <- data.frame(
        Test = c("Anderson-Darling Test", "Jarque-Bera Test", "Lilliefors Test", "Skewness", "Kurtosis"),
        `Test Statistic` = c(round(ad_result$statistic, 3),
                             round(jb_result$statistic, 3),
                             round(lillie_result$statistic, 3),
                             round(skew, 2),
                             round(kurt, 2)),
        `P-Value` = c(round(ad_result$p.value, 3),
                      round(jb_result$p.value, 3),
                      round(lillie_result$p.value, 3),
                      "N/A", "N/A"),
        Interpretation = c(
          ifelse(ad_result$p.value > 0.05, "Normal", "Non-Normal"),
          ifelse(jb_result$p.value > 0.05, "Normal", "Non-Normal"),
          ifelse(lillie_result$p.value > 0.05, "Normal", "Non-Normal"),
          ifelse(abs(skew) < 0.5, "Approximately Symmetric", ifelse(skew > 0, "Right-Skewed", "Left-Skewed")),
          ifelse(abs(kurt - 3) < 0.5, "Mesokurtic", ifelse(kurt > 3, "Leptokurtic (Heavy Tails)", "Platykurtic (Light Tails)"))
        )
      )
      kable(results, align = 'c') %>%
        kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
        add_header_above(c("Normality Test Summary" = 4))
    }

  # Time Series Summary Table
ts_summary_table <- function(ts_list) {
  
  ts_names <- names(ts_list)
  if (is.null(ts_names)) {
    ts_names <- paste("Series", seq_along(ts_list))  }
  
  summary_df <- data.frame(Series = character(),
                           Start = character(),
                           End = character(),
                           Min = numeric(),
                           Max = numeric(),
                           Mean = numeric(),
                           Median = numeric(),
                           stringsAsFactors = FALSE)

  for (i in seq_along(ts_list)) {
    ts_data <- ts_list[[i]]
    summary_df[i, ] <- c(
      ts_names[i],
      start(ts_data) %>% paste(collapse = "-"),
      end(ts_data) %>% paste(collapse = "-"),
      round(min(ts_data, na.rm = TRUE), 2),
      round(max(ts_data, na.rm = TRUE), 2),
      round(mean(ts_data, na.rm = TRUE), 2),
      round(median(ts_data, na.rm = TRUE), 2)
    )
  }

  kable(summary_df, align = 'c', 
        col.names = c("Series", "Start", "End", "Min", "Max", "Mean", "Median")) %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover", "condensed")) %>%
    add_header_above(c("Time Series Summary" = 7))
}


calculate_r2 <- function(predicted_values, actual_values) {
  ss_residual <- sum((actual_values - predicted_values)^2)
  ss_total <- sum((actual_values - mean(actual_values))^2)
  r_squared <- 1 - (ss_residual / ss_total)
  
  cat("R-squared of the model is:", round(r_squared, 4), "\n")}

ts_plots <- function(data, title = "Time Series Plots") {
  par(mfrow = c(4, 2))
  for (series_name in colnames(data)) {
    tsplot(data[, series_name], main = series_name, ylab = "Values", xlab = "Time")
  }
  title(main = title, outer = TRUE)
}


stationarity_summary <- function(data) {
  summary_table <- data.frame(
    Series = character(),
    `ADF Test` = character(),
    `Phillips-Perron Test` = character(),
    `KPSS Test` = character(),
    `ERS Test` = character(),
    stringsAsFactors = FALSE
  )

  for (series_name in colnames(data)) {
    series <- data[,series_name]

    adf_test <- adf.test(series, alternative = "stationary")
    pp_test <- pp.test(series)
    kpss_test <- kpss.test(series, null = "Level")
    urers_test <- ur.ers(series, type = "DF-GLS", model = "constant")

    adf_result <- ifelse(adf_test$p.value < 0.05, "Stationary", "Non-Stationary")
    pp_result <- ifelse(pp_test$p.value < 0.05, "Stationary", "Non-Stationary")
    kpss_result <- ifelse(kpss_test$p.value < 0.05, "Non-Stationary", "Stationary")
    urers_result <- ifelse(urers_test@teststat < urers_test@cval[2], "Stationary", "Non-Stationary")

    summary_table <- rbind(
      summary_table,
      data.frame(
        Series = series_name,
        `ADF Test` = adf_result,
        `Phillips-Perron Test` = pp_result,
        `KPSS Test` = kpss_result,
        `ERS Test` = urers_result,
        stringsAsFactors = FALSE
      )
    )
  }

  summary_table %>%
    kbl(align = 'c', col.names = c("Series", "ADF Test", "Phillips-Perron Test", "KPSS Test", "ERS Test")) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    add_header_above(c("Stationarity Test Interpretations" = 5))
}

garchAutoTryFit = function(
   ll,
   data,
   trace=FALSE,
   forecast.length=1,
   with.forecast=TRUE,
   ic="AIC",
   garch.model="garch" )
{
   formula = as.formula( paste( sep="",
                                "~ arma(", ll$order[1], ",", ll$order[2], ")+",
                                garch.model,
                                "(", ll$order[3], ",", ll$order[4], ")" ) )
   fit = tryCatch( garchFit( formula=formula,
                             data=data,
                             trace=FALSE,
                             cond.dist=ll$dist ),
                   error=function( err ) TRUE,
                   warning=function( warn ) FALSE )

   pp = NULL

   if( !is.logical( fit ) ) {
      if( with.forecast ) {
         pp = tryCatch( predict( fit,
                                 n.ahead=forecast.length,
                                 doplot=FALSE ),
                        error=function( err ) FALSE,
                        warning=function( warn ) FALSE )
         if( is.logical( pp ) ) {
            fit = NULL
         }
      }
   } else {
      fit = NULL
   }

   if( trace ) {
      if( is.null( fit ) ) {
         cat( paste( sep="",
                     "   Analyzing (", ll$order[1], ",", ll$order[2],
                                    ",", ll$order[3], ",", ll$order[4], ") with ",
                                    ll$dist, " distribution done.",
                     "Bad model.\n" ) )
      } else {
         if( with.forecast ) {
            cat( paste( sep="",
                        "   Analyzing (", ll$order[1], ",", ll$order[2], ",",
                                          ll$order[3], ",", ll$order[4], ") with ",
                                          ll$dist, " distribution done.",
                        "Good model. ", ic, " = ", round(fit@fit$ics[[ic]],6),
                        ", forecast: ",
                        paste( collapse=",", round(pp[,1],4) ), "\n" ) )
         } else {
            cat( paste( sep="",
                        "   Analyzing (", ll[1], ",", ll[2], ",", ll[3], ",", ll[4], ") with ",
                                          ll$dist, " distribution done.",
                        "Good model. ", ic, " = ", round(fit@fit$ics[[ic]],6), "\n" ) )
         }
      }
   }

   return( fit )
}

garchAuto = function(
   xx,
   min.order=c(0,0,1,1),
   max.order=c(5,5,1,1),
   trace=FALSE,
   cond.dists="sged",
   with.forecast=TRUE,
   forecast.length=1,
   arma.sum=c(0,1e9),
   cores=1,
   ic="AIC",
   garch.model="garch" )
{
   require( fGarch )
   require( parallel )

   len = NROW( xx )

   models = list( )

   for( dist in cond.dists )
   for( p in min.order[1]:max.order[1] )
   for( q in min.order[2]:max.order[2] )
   for( r in min.order[3]:max.order[3] )
   for( s in min.order[4]:max.order[4] )
   {
      pq.sum = p + q
      if( pq.sum <= arma.sum[2] && pq.sum >= arma.sum[1] )
      {
         models[[length( models ) + 1]] = list( order=c( p, q, r, s ), dist=dist )
      }
   }

   res = mclapply( models,
                   garchAutoTryFit,
                   data=xx,
                   trace=trace,
                   ic=ic,
                   garch.model=garch.model,
                   forecast.length=forecast.length,
                   with.forecast=TRUE,
                   mc.cores=cores )

   best.fit = NULL

   best.ic = 1e9
   for( rr in res )
   {
      if( !is.null( rr ) )
      {
         current.ic = rr@fit$ics[[ic]]
         if( current.ic < best.ic )
         {
            best.ic = current.ic
            best.fit = rr
         }
      }
   }

   if( best.ic < 1e9 )
   {
      return( best.fit )
   }

   return( NULL )
}
```

# III. Results & Conclusions

## 1. Descriptive Analysis

```{r}

culer = c(rgb(.85,.30,.12,.6), rgb(.12,.65,.85,.6))
# Plot the time series 
price_plot <- ggplot(total, aes(x = Date)) +
  geom_line(aes(y = Wheat, color = "Wheat")) +
  geom_line(aes(y = Corn, color = "Corn")) +
  labs(title = "Time Series of Wheat and Corn Futures Prices",
       y = "Price ($)",
       x = "Year") +
  scale_color_manual(values = c("Wheat" = "coral", "Corn"= "cornflowerblue")) +
  theme_light()

volume_plot <- ggplot(total, aes(x = Date)) +
  geom_line(aes(y = wheat_vol, color = "Wheat Volume"), linewidth = .5) +
  geom_line(aes(y = corn_vol, color = "Corn Volume"), linewidth = .25) +
  labs(title = "Time Series of Wheat and Corn Futures Volumes",
       y = "Volume",
       x = "Year") +
  scale_color_manual(values = c("Wheat Volume" = "coral", "Corn Volume"= "cornflowerblue"))+
  scale_y_continuous(labels = scales::label_number(scale_cut = cut_short_scale())) +
  theme_light()

price_plot / volume_plot
```

```{r}


p <- ggplot(total, aes(x=Month, color = Year)) + 
  geom_line(aes(y = Wheat), linetype = 3, alpha = 0.7, label =" ") +
  geom_line(aes(y = Corn), linetype = 4, alpha = 0.7) +
  labs(x = "Month",
       y = "Price ($)",
       title = "Wheat and Corn Futures Prices by Month") +
  theme_minimal() + 
  scale_x_continuous(breaks=c(2,4, 6, 8,10,12)) +
  transition_states(Year) 

animate(p, renderer = gifski_renderer())
```

## 2. Exploration

```{r}

# "time series has no or less than 2 periods"

# plot(decompose(wheat.ts))
# plot(stl(wheat.ts))


```

```{r}
# Separating Training & Testing data
testdata <- tail(total, 30)
total <- head(total,6293)


# Constructing xts objects
total.ts <- xts(total, order.by = total$Date)
wheat.ts <- xts(total$Wheat, total$Date)
corn.ts <- xts(total$Corn, total$Date)


# Original Data 
xts_plots(wheat.ts,"Wheat Futures",20)
xts_plots(corn.ts,"Corn Futures",20)

# Differenced Plots
xts_plots(diff(wheat.ts)[-1],"Differenced Wheat Futures",20)
xts_plots(diff(corn.ts)[-1],"Differenced Corn Futures")


# Daily returns 
# One way to calculate the returns is to use diff(log(data)), which is commented below. 
# xts_plots(diff(log(wheat.ts))[-1],"Returns on Wheat Futures",25)

# However, for the context of financial data, we can also use dailyReturns() from "quantmod"
wheat.r <- dailyReturn(wheat.ts)
xts_plots(wheat.r,"Returns on Wheat Futures",25)
xts_plots(wheat.r^2,"Squared Returns on Wheat Futures",25)

corn.r <- dailyReturn(corn.ts)
xts_plots(corn.r,"Returns on Corn Futures",25)
xts_plots(corn.r^2,"Squared Returns on Corn Futures",25)

# charts.RollingPerformance(R = wheat.r)

auto.arima(wheat.r) # ARIMA(0,0,0) with zero mean
auto.arima(corn.r) # ARIMA(0,0,0) with zero mean


```

```{r}

culer = c(rgb(.85,.30,.12,.6), rgb(.12,.65,.85,.6))
astsa::tsplot(na.approx(ts_ts(wheat.ts)), col=culer[1], lwd=2, pch=20, ylab="Price ($)", main="Wheat & Corn Futures - Ksmooth")
lines(ksmooth(time(ts_ts(wheat.ts)), na.approx(ts_ts(wheat.ts)), "normal", bandwidth=1), lwd=2, col=culer[1]) 

lines(na.approx(ts_ts(corn.ts)), col=culer[2], lwd=2, pch=20)
lines(ksmooth(time(ts_ts(corn.ts)), na.approx(ts_ts(corn.ts)), "normal", bandwidth=1), lwd=2, col=culer[2]) 

legend("topleft", col=culer, lty=1, lwd=2, pch=20, legend=c("Wheat Futures", "Corn Futures"), bg="white")

```

```{r}





auto.arima(wheat.ts) # ARIMA(1,1,3)
wheat_113 <- Arima(wheat.ts, order=c(1,1,3))
summary(wheat_113)
checkresiduals(wheat_113)
acf2(wheat_113$residuals) 
check_normality_residuals(wheat_113)

wheat_010 <- Arima(wheat.ts, order=c(0,1,0))
summary(wheat_010)
checkresiduals(wheat_010)
acf2(wheat_010$residuals) 

wheat_110 <- Arima(wheat.ts, order=c(1,1,0))
summary(wheat_110)
checkresiduals(wheat_110)
acf2(wheat_110$residuals) 

wheat_310 <- Arima(wheat.ts, order=c(3,1,0))
summary(wheat_310)
checkresiduals(wheat_310)
acf2(wheat_310$residuals) 

```

```{r}

auto.arima(corn.ts) # ARIMA(0,1,3)
corn_013 <- Arima(corn.ts, order=c(0,1,3))
summary(corn_013)
checkresiduals(corn_013)
acf2(corn_013$residuals) 
check_normality_residuals(corn_013)

corn_010 <- Arima(corn.ts, order=c(0,1,0))
summary(corn_010)
checkresiduals(corn_010)
acf2(corn_010$residuals) 

corn_210 <- Arima(corn.ts, order=c(2,1,0))
summary(corn_210)
checkresiduals(corn_210)
acf2(corn_210$residuals) 

corn_213 <- Arima(corn.ts, order=c(2,1,3))
summary(corn_213)
checkresiduals(corn_213)
acf2(corn_213$residuals) 

```

## 4. Garch

```{r}

# Returns of Wheat
mod_specify = ugarchspec(mean.model = list(armaOrder=c(1,1)),variance.model = list(model = "sGARCH",garchOrder=c(1,1)),distribution.model='sstd')

wheat.garch = ugarchfit(data = wheat.r, spec = mod_specify, out.sample = 20)

wheat.garch 

plot(wheat.garch, which=1)


```

```{r}


# Function to compare GARCH models
garch_compare <- function(data, arma_order = c(0, 0), garch_order = c(1, 1)) {
  # Supported conditional distributions
  cond_dists <- c("std", "snorm", "ged", "sged", "sstd", "snig", "QMLE")
  
  # Create a list to store fitted models
  models <- list()
  
  # Fit each GARCH model with the specified conditional distributions
  for (dist in cond_dists) {
    model_name <- paste0("ARMA(", arma_order[1], ",", arma_order[2], ")-GARCH(", 
                        garch_order[1], ",", garch_order[2], ") with ", dist)
    
    models[[model_name]] <- tryCatch(
      garchFit(
        formula = as.formula(paste0("~arma(", arma_order[1], ",", arma_order[2], ")+garch(", 
                                    garch_order[1], ",", garch_order[2], ")")),
        data = data,
        cond.dist = dist,
        trace = FALSE
      ),
      error = function(e) NULL
    )
  }
  
  # Create a summary table for AIC and BIC
  summary_df <- data.frame(Model = character(), AIC = numeric(), BIC = numeric(),stringsAsFactors = FALSE)
  for (i in seq_along(models)) {
        model <- models[[i]]
        summary_df[i, ] <- c(
          names(models)[i],
          round(model@fit[["ics"]][["AIC"]], 3),
          round(model@fit[["ics"]][["BIC"]], 3)) }
  # Output the table
  kable(summary_df, align = 'c', col.names = c("Model", "AIC", "BIC")) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    add_header_above(c("GARCH Model Comparison" = 3)) 
}

```

```{r}
garch_compare(wheat.r, arma_order = c(1, 1), garch_order = c(1, 1))
```

ARMA(0,0)-GARCH(1,1) with sstd -5.105\
ARMA(1,1)-GARCH(1,1) with sstd -5.106 -5.097

```{r}

wheat_auto = garchAuto(wheat.r, cores=8, trace=TRUE)
summary(wheat_auto)



corn_auto = garchAuto(corn.r, cores=8, trace=TRUE)
summary(corn_auto)
```

```{r}
garch_compare(corn.r, arma_order = c(1, 1), garch_order = c(1, 1))
```

ARMA(0,0)-GARCH(1,1) with sstd -5.484 -5.477 ARMA(0,0)-GARCH(1,1) with std -5.483 -5.478

```{r}
# Returns of Corn
mod_specify = ugarchspec(mean.model = list(armaOrder=c(1,1)),variance.model = list(model = "sGARCH",garchOrder=c(1,1)),distribution.model='std')

corn.garch = ugarchfit(data = corn.r, spec = mod_specify, out.sample = 20)

corn.garch 

plot(corn.garch, which=1)
```

```{r}
mod_specify = ugarchspec(mean.model = list(armaOrder=c(1,1)),variance.model = list(model = "eGARCH",garchOrder=c(1,1)),distribution.model='std')

corn.egarch = ugarchfit(data = corn.r, spec = mod_specify, out.sample = 20)

corn.egarch 
```

```{r}
spec_good <- mvspec(wheat.ts, log = "yes", main = "Spectrum of Wheat Futures (Log)", col = "coral", lwd = 2)

spec_good <- mvspec(corn.ts, log = "yes", main = "Spectrum of Corn Futures (Log)", col = "blue", lwd = 2)
```

```{r}
spec_good <- mvspec(wheat.r, log = "y", main = "Spectrum of Wheat Futures Returns (Log)", col = "coral", lwd = 2)

spec_good <- mvspec(corn.r, log = "y", main = "Spectrum of Corn Futures Returns (Log)", col = "blue", lwd = 2)
```

# IV. Closing

# V. Appendix & References
